### 2023-06-22 tex/conc.tex
[ writing ] start on introduction and conclusion

@@

Is this a reusable library?

Fitch-style systems (can't do).

>>>>

In this thesis, I have developed a foundation for semiring-annotated calculi in
natural deduction style.

@@

\begin{itemize}
  \item Novel simultaneous substitution for linear type theories
  \item Generalise \citet{AACMM21} work to substructural logics
  \item Is this a reusable library? Yes, examples
  \item Fitch-style systems (can't do).
  \item Methodology for metatheory of new calculi
  \item Semirings have been a reasonably robust basis.
  \item Can the code be directly generalised?
    Abstract away the algebraic structure of contexts.
\end{itemize}

### 2023-06-22 tex/intro.tex
[ more ] thesis writing
[ writing ] renew generic syntax chapter
[ writing ] start on introduction and conclusion

%\begin{itemize}
%  \item Review standard linear algebra constructions in the posemiring setting.
%  \item Give a general definition of simple type systems with usage annotations
%    based on connectives from bunched logic and constructions from linear
%    algebra.
%  \item Give some novel examples of such systems.
%  \item Derive admissibility of renaming and substitution for all such systems.
%  \item Give a usage checker generic in syntax.
%\end{itemize}

@@

We build upon both the general structural framework of
\citet{AACMM21} and the substructural techniques of \citet{WA20}.
The way \citeauthor{AACMM21} consolidate and codify mechanisation techniques for
propositional natural deduction systems based on intrinsically typed syntax and
de Bruijn indices, we aim to replicate for linear-like systems based on
semiring usage annotations.
By picking a trivial semiring, our work can subsume that of
\citeauthor{AACMM21}, except for the many pieces of machinery we have not yet
ported to this new framework.

@@

This paper proceeds as follows.
In \cref{sec:algebra}, we review and fix conventions pertaining to partially
ordered semirings and vectors over them.
In \cref{sec:bunched-rules}, we introduce an informal meta-syntax allowing us
to state substructural typing rules succinctly and without explicit reference
to contexts.
In \cref{sec:syntax}, we mechanise that meta-syntax, giving a type of
\emph{descriptions} of type systems, and interpreting those descriptions as
types of intrinsically typed terms.
In \cref{sec:env}, we discuss usage-aware environments: a generalisation of
the structures used in simultaneous renaming and substitution proofs.
We use environments in \cref{sec:semantics} to state an alternative
elimination principle for terms, and give examples of such eliminations in
\cref{sec:examples}.
The examples are syntax-generic renaming and substitution, a specific
denotational semantics, and a syntax-generic usage elaborator.
Finally, we conclude and discuss future work in \cref{sec:conc}

The work presented in this paper has been mechanised in Agda,
with the code available for building upon~\cite{generic-lr}.

>>>>

\chapter{Introduction}

FIXME: the start of this chapter is directly from the ESOP22 paper.

@@

We build upon both the general structural framework of
\citet{AACMM21} and the substructural techniques of \citet{WA21}.
The way \citeauthor{AACMM21} consolidate and codify mechanisation techniques for
propositional natural deduction systems based on intrinsically typed syntax and
de Bruijn indices, we aim to replicate for linear-like systems based on
semiring usage annotations.
By picking a trivial semiring, our work can subsume that of
\citeauthor{AACMM21}, except for the many pieces of machinery we have not yet
ported to this new framework.

@@

\section{Naming and notation conventions}

I assume familiarity with the Curry-Howard correspondence throughout this
thesis.
I make no distinction between logics and type theories, and use terminology from
each interchangeably.
Each following bullet point lists a collection of synonyms.

\begin{itemize}
  \item assumption, hypothesis, variable
  \item derivation, proof, term
  \item proposition, formula, type
  \item connective, type former
\end{itemize}

I carry out mechanised constructions and proofs in the proof assistant and
programming language Agda~\citep{Agda}.
Agda is based on Martin-L\"{o}f's intensional dependent type theory, so I
similarly present non-mechanised constructions and proofs assuming a foundation
given by dependent type theory, in a style inspired by the HoTT
Book~\citep{hottbook}.

\section{Outline of the thesis}

This thesis proceeds as follows.
The next two chapters, \cref{sec:simple,sec:linearity}, are introductory in
nature, and cover two largely independent strands of prior work.
In \cref{sec:simple}, I introduce existing methods of representing and reasoning
about type systems in proof assistants based on dependent type theory.
I start from well established representations of well scoped and well typed
terms, and develop these towards a recent approach to environment-based
semantics given by \citet{AACMM21}.
In \cref{sec:linearity}, I discuss the challenges faced when one extends a
treatment of a simple type system, such as that given in \cref{sec:simple}, to
modal and linear type systems.
We see that modal and linear type systems apparently violate some of the nice
properties of the simply typed $\lambda$-calculus we required in
\cref{sec:simple}.
I present a solution for intuitionistic S4 modal logic, but leave a solution for
linear logic to the following chapters.

In the following two chapters, \cref{sec:semirings,sec:ren-sub-lr}, I present a
calculus $\name$ parametrised by a partially ordered semiring of \emph{usage
annotations}.
In \cref{sec:semirings}, I define the calculus, give some possible extensions,
and show that it subsumes intuitionistic S4 modal logic and Intuitionistic
Linear Logic.
In \cref{sec:ren-sub-lr}, I show that $\name$ enjoys generalised versions of the
nice properties required in \cref{sec:simple}, and I proceed to give novel
definitions of simultaneous substitutions and their action on $\name$ terms.
These two chapters are adapted from the work of \citet{WA21}.

The remaining three main chapters,
\cref{sec:framework,sec:semantics,sec:example-semantics}, adapt the syntactic
and semantic framework of \citet{AACMM21}, as presented at the end of
\cref{sec:simple}, to semiring-annotated calculi.
\Cref{sec:framework,sec:semantics} generalise the work on $\name$ presented in
\cref{sec:semirings,sec:ren-sub-lr}, respectively.
\Cref{sec:framework} shows how to formally describe the syntax of an arbitrary
semiring-annotated calculus, following the constructions used in
\cref{sec:semirings}.
\Cref{sec:semantics} then provides the generic environment-based semantic
traversal on such syntaxes, providing renaming and substitution as per
\cref{sec:ren-sub-lr} for all syntaxes as special cases of the generic
traversal.
\Cref{sec:examples} then gives further example uses of the generic traversal.

Finally, I conclude with \cref{sec:conc}, which discusses the achievements of
this thesis and openings for future work.
